sentiment-analysis

### 2016-08-06 eda/sentiment-analysis.Rmd
formatting and analysis
analysis

@@

```{r, eval = T}
# boxplot for each group
ggplot(review_scores) +
  geom_boxplot(aes(stars, sentiment, group = stars)) +
  labs(x = "Business Star Rating",
       y = "Mean Sentiment Score",
       title = "Mean Sentiment Score vs. Business Star Rating") +
  theme_bw()
```

@@

```{r fig.height = 7, fig.width = 7}
summaries %>% 
  ggplot(aes(reviews, average_star, color = score)) +
  geom_point() +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1, hjust = 1) +
  scale_x_log10(limits = c(200, 100000), 
                breaks = c(1000, 10000), labels = c(1000, 10000)) +
  geom_hline(yintercept = mean(reviews$stars), color = "black", lty = 2) +
  scale_color_gradient2(low = "red", high = "blue", 
                        midpoint = 0, mid = "grey") +
  labs(x = "# of reviews",
       y = "Average Yelp Stars",
       color = "Valence Score") +
  scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, 1)) +
  theme_bw()
```

@@

```{r fig.height = 7}
summaries %>% 
  arrange(desc(reviews)) %>% 
  ggplot(aes(score, average_star)) +
  geom_point(aes(size = reviews)) +
  geom_text(aes(label = word), vjust = 1, hjust = 1, check_overlap = TRUE) +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Valence Score") +
  ylab("Average Yelp Stars") +
  labs(title = "Average Yelp Star Rating vs. Valence Score") +
  scale_x_continuous(limits = c(-5,6), breaks = seq(-5, 5, 2)) +
  scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, 1)) +
  expand_limits(x = -6)
```

>>>>

write_csv(reviews_text, "../clean_data/reviews_text.csv")

@@

```{r, eval = T}
# boxplot for each group
ggplot(review_scores) +
  geom_boxplot(aes(stars, sentiment, group = stars)) +
  labs(x = "Business Star Rating",
       y = "Mean Sentiment Score",
       title = "Mean Sentiment Score vs. Business Star Rating")
```

@@

```{r fig.height = 7, fig.width = 7}
summaries %>% 
  ggplot(aes(reviews, average_star, color = score)) +
  geom_point() +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1, hjust = 1) +
  scale_x_log10(limits = c(200, 100000), 
                breaks = c(1000, 10000), labels = c(1000, 10000)) +
  geom_hline(yintercept = mean(reviews$stars), color = "black", lty = 2) +
  scale_color_gradient2(low = "red", high = "blue", 
                        midpoint = 0, mid = "grey") +
  labs(x = "# of reviews",
       y = "Average Yelp Stars",
       color = "Valence Score") +
  scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, 1))
```

@@

This is pretty informative, but I want to see these words plotted by their valence score and average Yelp rating to visualize if there's a positive correlation here between sentiment score and their star rating. 
```{r fig.height = 7}
summaries %>% 
  arrange(desc(reviews)) %>% 
  ggplot(aes(score, average_star)) +
  geom_point(aes(size = reviews)) +
  geom_text(aes(label = word), vjust = 1, hjust = 1, check_overlap = TRUE) +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Valence Score") +
  ylab("Average Yelp Stars") +
  labs(title = "Average Yelp Star Rating vs. Valence Score") +
  scale_x_continuous(limits = c(-5,6), breaks = seq(-5, 5, 2)) +
  scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, 1)) +
  expand_limits(x = -6)
```

Based on a simple linear regression line, words with higher scores are clustered towards 4 and 5-star ratings, while most negative words have pretty low average ratings. Some words like "damn" had higher star ratings despite their low AFINN score, but that could be explained by reviews where writers could have described their food or service as "damn good".
